---
title: "P2"
author: "Wang Luyuan"
date: "2019/8/24"
output: html_document
---
# Overview:
This report analyzes the `ToothGrowth` data in the R data sets package.The full code is pulished on (https://github.com/Wang-CODEPROJECT/R-statstical-inference)

```{r}
setwd('D:/R-Studio/Coursera/Stats/Statistical-Inference-master/part 2')
suppressWarnings(library(ggplot2)); suppressMessages(library(data.table)); 
library(grid);
source('multiplot.R');
```

### Data Cleansing:
From the source, we will load the `ToothGrowth` data into a `data.table` object, change the column names to something more meaningful, and declare a join key.

```{r}
# load data and make column names meaningful
dt<-data.table(ToothGrowth)
setnames(dt,c('len','supp','dose'),c('Length','Supplement','Dose'))

# add 'Dosage'and set the join key
dt<-dt[,Dosage:=sapply(as.character(dt$Dose),function(x) as.factor(switch(x,'0.5'='SM','1'='MD','2'='LG')))]
setkey(dt,Supplement,Dosage)
head(dt,1)
```

### Exploratory Analysis:
The following result sets are two a simple exploratory methods to understand the content and the structure of the `data.table` 
```{r}
summary(dt)
```

```{r}
str(dt)
```

To further conduct the exploratory analysis, we can plot `Length` against both `Dosage` and `Supplement`.  When we do this we see that the larger the `Dosage`, the longer the tooth `Length`.  However, it is slightly unclear as to which supplement is more effective, Orange Juice `OJ` or Ascorbic Acid `VC`.

```{r}
# plot 1 
g1<-ggplot(dt,aes(x=Dosage,y=Length))
g1<-g1+geom_point(aes(color=Supplement),size=4)
# plot 2
g2<-ggplot(dt,aes(x=Supplement,y=Length))
g2<-g2+geom_point(aes(color=Dosage),size=4)
# plot together
multiplot(g1,g2,cols=2)
```

#### *Compare Dosage Alone*

```{r}
t1<-subset(dt,Dosage=='SM')$Length
t2<-subset(dt,Dosage=='MD')$Length
t<-t.test(t1,t2,paired=FALSE,var.equal=FALSE)
t$conf.int[1:2]
```

We increase the Vitamin C dose from 0.5 to 1.0 milligrams, the confidence interval does not contain zero, so we can reject the null hypothesis that this dose increase does not increase tooth length.

```{r}
t1<-subset(dt,Dosage=='MD')$Length
t2<-subset(dt,Dosage=='LG')$Length
t<-t.test(t1,t2,paired=FALSE,var.equal=FALSE)
t$conf.int[1:2]
```

Next, if we increase the Vitamin C dose from 1.0 to 2.0 milligrams, the confidence interval againg does not contain zero, so we can reject the null hypothesis that this dose increase does not increase tooth length.

#### *Compare Supplement Alone*
```{r}
t1<-subset(dt,Supplement=='VC')$Length
t2<-subset(dt,Supplement=='OJ')$Length
t<-t.test(t1,t2,paired=FALSE,var.equal=FALSE)
t$p.value
t$conf.int[1:2]
```

#### *Compare Supplement by Each Dosage*
``` {r}
t1<-subset(dt,Supplement=='VC' & Dosage=='SM')$Length
t2<-subset(dt,Supplement=='OJ' & Dosage=='SM')$Length
t<-t.test(t1,t2,paired=FALSE,var.equal=FALSE)
t$conf.int[1:2]
```

When we continue the analysis, and compare a 'SM' dosage of Ascorbic Acid to a'SM' dosage of Orange Juice, we see the confidence interval does not contain zero, so we can reject the null hypothesis that supplement type with a 'SM' dosage does not affect tooth growth.

```{r}
t1<-subset(dt,Supplement=='VC' & Dosage=='MD')$Length
t2<-subset(dt,Supplement=='OJ' & Dosage=='MD')$Length
t<-t.test(t1,t2,paired=FALSE,var.equal=FALSE)
t$conf.int[1:2]
```

Next, we compare a 'MD' dosage of Ascorbic Acid to a 'MD' dosage of Orange Juice, and, again, we see the confidence interval does not contain zero; so, we can reject the null hypothesis that supplement type with a 'MD' dosage does not affect tooth growth.

```{r}
t1<-subset(dt,Supplement=='VC' & Dosage=='LG')$Length
t2<-subset(dt,Supplement=='OJ' & Dosage=='LG')$Length
t<-t.test(t1,t2,paired=FALSE,var.equal=FALSE)
t$p.value
t$conf.int[1:2]
```

Lastly, we compare a 'LG' dosage of Ascorbic Acid to a 'LG' dosage of Orange Juice; this time, however, we observer the confidence interval contains zero and there is a p-value of almost 1.0.  In turn, we do not reject the null hypothesis that supplement type with a 'LG' dosage does not affect tooth growth.  Meaning, with a 'LG' Dosage, we cannot conclude which supplement type has a greater affect on tooth growth.
  
    
### Conclusions:
1. As Vitamin C dose size alone increases, the tooth length increases as well, and
2. Irrespective of dose size, supplement type alone does not affect tooth growth; however,
3. The supplement type of Orange Juice, or 'OJ', affects tooth length greater then Ascorbic Acid, or 'VC', with a 0.5 and 1.0 dose size, in turn, 
4. When the dose size reached 2.0 milligrams, there is no difference between Orange Juice and Ascorbic Acid.